---
title: "Homework 1 - Time Series Toolbox"
author: "Prof. Dr. Stephan Trahasch"
date: 'Submission date: 17.12.2020'
output:
  html_document:
    theme: cerulean
    css: styles/homework.css
    toc: yes
    toc_depth: 3
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '3'
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
options(digits = 5)

# libraries to load
library(fpp3)
library(fpp2)
```

# Exercise 1

For the following series, find an appropriate Box-Cox (?BoxCox) transformation in order to stabilize the variance. First plot the time series and decide which of them need a transformation. 

  * `usnetelec`
  * `usgdp`
  * `mcopper`
  * `enplanements`

```{r}
# todo
autoplot(usnetelec)
autoplot(usgdp)
autoplot(mcopper)
autoplot(enplanements)
```


```{r}
help("BoxCox") 
lambdaCopper = BoxCox.lambda(mcopper)
autoplot(BoxCox(mcopper, lambdaCopper))

lambdaEnplanements = BoxCox.lambda(enplanements)
autoplot(BoxCox(enplanements, lambdaEnplanements))
```


# Exercise 2

Why is a Box-Cox transformation unhelpful for the `cangas` data?

```{r}
# todo
 autoplot(cangas)

```
Your answer: Eine BoxCox transformation ist hauptsächlich dann hilfreich, wenn die Varianz über die Zeit steigt. Der Datensatz Cangas hat aber, wie hier deutlich zu sehen ist eine bauchige Form, wodurch die Varianz lediglich in der Mitte hoch ist, am Anfang und Ende aber klein.


# Exercise 3

What Box-Cox transformation would you select for **your** retail data from Exercise 4 in Visualization?

```{r}
# Example 
retaildata <- readxl::read_excel("data/retail.xlsx", skip=1)
myts <- ts(retaildata[,"A3349873A"], frequency=12, start=c(1982,4))
autoplot(myts)
myts %>% BoxCox(lambda = 0) %>% autoplot()
```


From visual inspection, a log transformation would be appropriate here. It also makes sense, as retail expenditure is likely to increase proportionally to population, and therefore the seasonal fluctuations are likely to be proportional to the level of the series. It has the added advantage of being easier to explain than some other transformations. Finally, it is relatively close to the automatically selected value of `BoxCox.lambda(myts)` $= `r round(BoxCox.lambda(myts),3)`$.
```{r}
retaildata <- readxl::read_excel("data/retail.xlsx", skip=1)
myts <- ts(retaildata[,"A3349335T"], frequency=12, start=c(1982,4))
autoplot(myts)
lambdaSelf = BoxCox.lambda(myts)
print(lambdaSelf)
myts %>% BoxCox(lambda = 0) %>% autoplot()
```
If you have selected a different series from the retail data set, you might choose a different transformation.

Your answer: In meinem Beispiel ist ebenfalls eine logarithmische Transformation sinnvoll. Man sieht gut, wie die Varianz bei der transformierten Abbildung kleiner ist. Zudem gibt der automatisch ausgewählte lambda-wert für meine Funktion ebenfalls mit 0.19385 einen Wert nahe bei 0 zurück.


# Exercise 4

Calculate the residuals (?residuals) from a seasonal naive forecast applied to the quarterly Australian beer production data from 1992. The following code will help.

```{r}
# todo
(?residuals)
beer_fc = snaive(ausbeer, h = 40)
res = residuals(beer_fc)

autoplot(ausbeer) + autolayer(beer_fc)

```



Test if the residuals are white noise and normally distributed.

```{r}
checkresiduals(beer_fc)
```

What do you conclude?

Your answer: Die Residuals sind laut des Correlograms nicht nur white noise. Das ist nicht so gut, da das für ein nicht so gutes Vorhersage-Modell steht. 


# Exercise 5

Are the following statements true or false? Explain your answer.

> a. Good forecast methods should have normally distributed residuals.

Your answer: wahr, da es das Berechnen von Vorhersage-Intervallen erleichtert.

> b. A model with small residuals will give good forecasts.

Your answer wahr. Residuals sind die Abweichungen der Vorhersagen zu den tatsächlichen Werten. Das bedeutet, je kleiner die Residuals sind, desto besser ist auch die Vorhersagemethode.

> c. The best measure of forecast accuracy is MAPE.

Your answer: falsch, das beste Vorhersage Model findet man am besten anhand des geringsten RMSE Wert.

> d. If your model doesn't forecast well, you should make it more complicated.

Your answer: falsch, man sollte eher versuchen, den Fehler zu finden, wieso es nicht gut vorhersagt. Eventuell kann man die Daten in der Vorverarbeitung so vorbereiten, dass das Model bessere Vorhersagen trifft. 

> e. Always choose the model with the best forecast accuracy as measured on the test set.

Your answer: wahr, falls die Test-Daten die gleichen Eigenschaften haben wie die origninal-Daten. Dann kann man davon ausgehen, dass das Model auch gut auf den original-Daten performt.

# Exercise 6

For your retail time series (from Exercise 4):

Split the data into two parts using

```{r}
retaildata <- readxl::read_excel("data/retail.xlsx", skip=1)
myts <- ts(retaildata[,"A3349335T"], frequency=12, start=c(1982,4))
myts.train <- window(myts, end=c(2010,12))
myts.test <- window(myts, start=2011)
```

Check that your data have been split appropriately by producing the following plot.

```{r}
autoplot(myts) +
  autolayer(myts.train, series="Training") +
  autolayer(myts.test, series="Test")
```

Calculate forecasts using `snaive` applied to `myts.train`.

```{r}
# todo
forecast = snaive(myts.train)

forecast_test = snaive(myts.test)
```

Compare the accuracy of your forecasts against the actual values stored in `myts.test`.
(?accuracy) 

```{r}
# todo
(?accuracy)
accuracy(forecast)
 
accuracy(forecast_test)
```

The number to look at here is the test set RMSE of 71.443. That provides a benchmark for comparison when we try other models.

Check the residuals. Do the residuals appear to be uncorrelated and normally distributed?

```{r}
# todo
 checkresiduals(forecast)
```

Your answer: Nein, die Residuals sehen nicht unkorreliert aus. An dem Correlogram erkennt man, dass die Residuals stark nach oben ausschlagen, also eine positive Korrelation vorweisen. Die Residuals sehen allerdings nahezu Normalverteilt aus.


How sensitive are the accuracy measures to the training/test split?
lediglich an dem MPE und MAPE measure, sieht man größere Differenzen zwischen den trainings und den test Daten.


# Exercise 7

`visnights` contains quarterly visitor nights (in millions) from 1998-2015 for eight regions of Australia.

Use `window()` to create three training sets for `visnights[,"QLDMetro"],` omitting the last 1, 2 and 3 years; call these train1, train2, and train3, respectively.

```{r}
train1 <- window(visnights[, "QLDMetro"], end = c(2015, 3))
train2 <- window(visnights[, "QLDMetro"], end = c(2015, 2))
train3 <- window(visnights[, "QLDMetro"], end = c(2015, 1))
```

Compute one year of forecasts for each training set using the `snaive()` method. Call these `fc1`, `fc2` and `fc3`, respectively.

```{r}
# todo
fc1 = snaive(train1)
fc2 = snaive(train2)
fc3 = snaive(train3)
```

Use `accuracy()` to compare the MAPE over the three test sets. Comment on these.
First we will copy the actual data into a variable. Then we can do an accuracy comparison.

```{r}
qld <- visnights[, "QLDMetro"]
# todo
accuracy(fc1)
accuracy(fc2)
accuracy(fc3)
```

This should give similar results to this consolidated results table.

```
                     ME     RMSE      MAE        MPE     MAPE      MASE       ACF1
Training set  0.1637836 1.742687 1.360271  0.4384347 7.357322 1.0000000 0.06643175
Test set fc1 -1.3010774 1.301077 1.301077 -6.9956861 6.995686 0.9564839         NA
Test set fc2 0.08383478 1.387447 1.384912 -0.4063445 6.589342 1.019346 -0.50000000
Test set fc3 0.06202858 1.132896 0.9294135 -0.237857 4.425934 0.6738562 -0.51548610
```

The lower MAPE value for "fc3" indicates a better result when we use the previous 3 values for the `snaive()` prediction.

# Exercise 8

Use the Dow Jones index (data set `dowjones`) to do the following:

Produce a time plot of the series.

```{r}
# todo
help(downjones)
```

Produce forecasts using the drift method and plot them.

Let's assume we want to forecast the next 5, 10 and 15 values.

```{r}
dowfc1 <- rwf(dowjones, drift=TRUE, h=5)
# todo
dowfc1 <- rwf(dowjones, drift=TRUE, h=5)
dowfc2 <- rwf(dowjones, drift=TRUE, h=10)
dowfc3 <- rwf(dowjones, drift=TRUE, h=15)
```
```{r}
autoplot(dowfc1) 
autoplot(dowfc2)
autoplot(dowfc3)
```

Then we can plot these values.

```{r eval = FALSE}
# remove eval=FALSE
autoplot(dowjones) +
  autolayer(dowfc1, PI=FALSE, series="Drift 5") +
  autolayer(dowfc2, PI=FALSE, series="Drift 10") +
  autolayer(dowfc3, PI=FALSE, series="Drift 15") +
  xlab("Time") + ylab("Closing Price (US$)") +
  ggtitle("Dow Jones index") +
  guides(colour=guide_legend(title="Forecast"))
```

We show that the forecasts are identical to extending the line drawn between the first and last observations.

We can plot the forecasts in a different order, so the shorter forecasts are superimposed, showing the lines are the same.

```{r eval = FALSE}
# remove eval=FALSE
autoplot(dowjones) +
  autolayer(dowfc3, PI=FALSE, series="Drift 15") +
  autolayer(dowfc2, PI=FALSE, series="Drift 10") +
  autolayer(dowfc1, PI=FALSE, series="Drift 5") +
  xlab("Time") + ylab("Closing Price (US$)") +
  ggtitle("Dow Jones index") +
  guides(colour=guide_legend(title="Forecast"))
```

Try using some of the other benchmark functions to forecast the same data set. Which do you think is best? Why?

The time series isn't seasonal, so the seasonal naive method is not viable. However, we can use the mean and naive methods.

```{r}
# todo
fcmean <- meanf(dowjones)
fcnaive <- naive(dowjones)
```
```{r}
autoplot(fcnaive)
autoplot(fcmean)
```
```{r}
accuracy(fcmean)
accuracy(fcnaive)
```

The three values will be very different here. The Mean will use the data set, so is unlikely to follow the current trendline.
Aus diesem Grund ist hier die Naive Methode die beste.

# Exercise 9

Consider the daily closing IBM stock prices (data set `ibmclose`).

Produce some plots of the data in order to become familiar with it.

```{r}
# todo
autoplot(ibmclose)
ggAcf(ibmclose)
gglagplot(ibmclose)
```

Split the data into a training set of 300 observations and a test set of 69 observations.

```{r}
# todo
ibm.train <- window(ibmclose, end=300)
ibm.test <- window(ibmclose, start=301)
```

Try using various benchmark methods to forecast the training set and compare the results on the test set. Which method did best?
Drift Methode.

```{r eval = FALSE}
# remove eval=FALSE
h <- length(ibm.test)
m.f <- meanf(ibm.train, h=h)
# todo
rw.f <- naive(ibm.train, h=h)
rwd.f <- rwf(ibm.train, drift=TRUE, h=h)


autoplot(ibmclose) +
  xlab("Day") +
  ggtitle("Daily closing IBM stock prices") +
  autolayer(m.f$mean, col=2, series="Mean method") +
  autolayer(rw.f$mean, col=3, series="Naive method") +
  autolayer(rwd.f$mean, col=4, series="Drift method")

# accuracy(m.f,ibm.test)
# todo
accuracy(m.f, ibm.test)
accuracy(rw.f, ibm.test)
accuracy(rwd.f, ibm.test)
```


Check the residuals of your preferred method. Do they resemble white noise?
Bis auf drei Ausreißer, ähnelt es white noise schon ein bisschen.

```{r}
# todo
 checkresiduals(rwd.f)
```


# Exercise 10

Consider the sales of new one-family houses in the USA, Jan 1973 -- Nov 1995 (data set `hsales`).

Produce some plots of the data in order to become familiar with it.

```{r}
# todo
print(hsales)
autoplot(hsales)
ggAcf(hsales)
ggsubseriesplot(hsales)
ggseasonplot(hsales)
gglagplot(hsales)
```

Split the `hsales` data set into a training set and a test set, where the test set is the last two years of data.

```{r}
# todo
sales.train <- window(hsales, end=1993)
sales.test <- window(hsales, start=1994)
```
```{r}
autoplot(sales.test)
```

Try using various benchmark methods to forecast the training set and compare the results on the test set. Which method did best?

```{r}
# todo
h <- length(sales.test)
salesMean.f <- meanf(sales.train, h=h)
salesNaive.f <- naive(sales.train, h=h)
salesSnaive.f <- snaive(sales.train, h=h)
salesDrift.f <- rwf(sales.train, drift=TRUE, h=h)
```
```{r}
accuracy(salesMean.f, sales.test)
accuracy(salesNaive.f, sales.test)
accuracy(salesDrift.f, sales.test)
accuracy(salesSnaive.f, sales.test)
```

In terms of accuracy measures on the test set, the seasonal naive method does better.
In meinem Fall ist die Accuracy für die Mean Methode am besten.

Check the residuals of your preferred method. Do they resemble white noise?

```{r eval = FALSE}
# remove eval=FALSE
checkresiduals(salesSnaive.f)
checkresiduals(salesMean.f)
```
Die Residuen sehen hier in diesem Fall nicht nach white noise aus. Es gibt eine hohe positive Korrelation zu beginn und gegen Ende eine leichte negative Korrelation.

