---
title: "Homework 1 - Time Series Visualization"
author: "Prof. Dr. Stephan Trahasch"
date: 'Submission date: 17.12.2020'
output:
  html_document:
    theme: cerulean
    css: styles/homework.css
    toc: yes
    toc_depth: 3
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '3'
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
options(digits = 5)

# libraries to load
library(fpp3)
library(fpp2) 
```

# Exercise 1
Use the help menu to explore what the series `gold`, `woolyrnq` and `gas` represent. These are available in the `forecast` package.

```{r, echo=TRUE, cache=TRUE, eval=FALSE}
help(gold)
# to do
```
#Gold help
gold {forecast}	R Documentation
Daily morning gold prices
Description
Daily morning gold prices in US dollars. 1 January 1985 – 31 March 1989.

Usage
gold
Format
Time series data

```{r}
help(woolyrnq)
```
#Woolyrnq help
woolyrnq {forecast}	R Documentation
Quarterly production of woollen yarn in Australia
Description
Quarterly production of woollen yarn in Australia: tonnes. Mar 1965 – Sep 1994.

Usage
woolyrnq
Format
Time series data

Source
Time Series Data Library. https://pkg.yangzhuoranyang.com/tsdl/
```{r}
help(gas)
```
#Gas help
gas {forecast}	R Documentation
Australian monthly gas production
Description
Australian monthly gas production: 1956–1995.

Usage
gas
Format
Time series data

Source
Australian Bureau of Statistics.
________________________________________________________
Use `autoplot` to plot each of these in separate plots.

```{r, echo=TRUE, cache=TRUE}
# to do
autoplot(gold)
autoplot(woolyrnq)
autoplot(gas)

```

**What is the frequency of each commodity series?**
Hint: apply the `frequency()` function.

```{r, echo=TRUE, cache=TRUE}
# to do
frequency(gold)
frequency(woolyrnq)
frequency(gas)
```

Use `which.max()` to spot the outlier in the `gold` series. 
**Which observation was it?**

```{r, echo=TRUE, cache=TRUE}
# to do
which.max(gold)
gold[770]
```
It was observation 770 with a value of 593.7

# Exercise 2 

Import the file `tute1.csv` from the folder data and review its contents. You should find four columns of information. Columns B through D each contain a quarterly series, labelled Sales, AdBudget and GDP. Sales contains the quarterly sales for a small company over the period 1981-2005. AdBudget is the advertising budget and GDP is the gross domestic product. All series have been adjusted for inflation.

You can read the data into R with the following script:

```{r, echo=TRUE, cache=TRUE}
tute1 <- read.csv("data/tute1.csv", header = TRUE)
View(tute1)
```

We convert the data to time series. The `[,-1]` removes the first column which contains the quarters as we don't need them now.

```{r, echo=TRUE, cache=TRUE}
mytimeseries <- ts(tute1[, -1], start = 1981, frequency = 4)
```

Construct time series plots of each of the three series and use 'facets = TRUE' .

```{r, echo=TRUE, cache=TRUE}
# to do

autoplot(mytimeseries, facets = TRUE)
```

Check what happens when you don't include `facets=TRUE`.

```{r, echo=TRUE, cache=TRUE, dependson="exmyts"}
# to do
autoplot(mytimeseries)
```

# Exercise 3

Import some monthly Australian retail data from the folder data. These represent retail sales in various categories for different Australian states, and are stored in a MS-Excel file.

You can read the data into R with the following script:

```{r, echo=TRUE, cache=TRUE}
retaildata <- readxl::read_excel("data/retail.xlsx", skip = 1)
```

The second argument (`skip=1`) is required because the Excel sheet has two header rows.

Select one of the time series as follows (but replace the column name with your own chosen column):

```{r, echo=TRUE, cache=TRUE}
myts <- ts(retaildata[, "A3349335T"], frequency = 12, start = c(1982, 4))
```

Explore your chosen retail time series using the following functions: `autoplot`, `ggseasonplot`, `ggsubseriesplot`, `gglagplot`, `ggAcf`

```{r, echo=TRUE, cache=TRUE}
# to do
autoplot(myts)
ggseasonplot(myts)
ggsubseriesplot(myts)
gglagplot(myts)
ggAcf(myts)
```

**Can you spot any seasonality, cyclicity and trend? What do you learn about the series?**

Your answer: 
The Trend is going up over a long time. It is clearly visible in the autoplot.
In the ggseasonplot it is visible, that every year the values go up aswell.
The subseriesplot shows, that there is a seasonality. The Values for December are higher as the other values for every year.




# Exercise 4
Create time plots of the following time series: goog, writing, fancy, a10, h02.

    + Use `help()` to find out about the data in each series.
    + For the `goog` plot, modify the axis labels and title.

```{r, echo=TRUE, message=FALSE, warning=FALSE}
# to do
help(goog)
autoplot(goog, xlab="Trading days", ylab="Stock prices", main="Daily closing stock prices of Google Inc")
autoplot(writing)
autoplot(fancy)
autoplot(a10)
autoplot(h02)

```

Use the `ggseasonplot` and `ggsubseriesplot` functions to explore the seasonal patterns in the following time series:  `writing`, `fancy`, `a10`, `h02`.

- What can you say about the seasonal patterns?
- Can you identify any unusual years?


```{r, echo=TRUE}
# writing
# to do
help(writing)
ggseasonplot(writing)
ggsubseriesplot(writing)

```

Your answer:
There is a seasonal pattern, that shows probably the summer holidays in school. Therefore less Paper for printing and wirting is bought.


```{r, echo=TRUE}
# fancy
# to do
help(fancy)
ggseasonplot(fancy)
ggsubseriesplot(fancy)

```

Your answer:
In general there is a trend going up from year to year.
The Seasonality is clearly visible for december. The Souvenir shops make a lot of sales, probably because a lot of people go to Australia over winter.


```{r, echo=TRUE}
# a10
# to do
help(a10)
ggseasonplot(a10)
ggsubseriesplot(a10)

```

Your answer:
One can see a seasonality for January. Maybe the Christmas Time brings a lot of Sweets to the people, which could be the reason why the diabetics need insulin as the anti-diabetic drug.

The year 2008 is a little different because there is no data from june on. 
But there is also a slower decrease after january than in other years.


```{r, echo=TRUE}
# h02
# to do
help(h02)
ggseasonplot(h02)
ggsubseriesplot(h02)
```

Your answer:
There is a clearly visible Trend that each year more corticosteroid drugs are used.
The seasonality is also very clear. in February there is a big decrease in corticosteroid drugs use.
2008 is again a little different, because the decrease after january is slower.



# Exercise 5
Use the following graphics functions: `autoplot`, `ggseasonplot`, `ggsubseriesplot`, `gglagplot`, `ggAcf` and explore features from the following time series: `hsales`, `usdeaths`, `bricksq`, `sunspotarea`, `gasoline`.

    + Can you spot any seasonality, cyclicity and trend?
    + What do you learn about the series?

Example for hsales:

```{r, echo=TRUE, cache=TRUE}
autoplot(hsales)
ggseasonplot(hsales)
ggsubseriesplot(hsales)
gglagplot(hsales)
ggAcf(hsales)
```

  + Seasonality evident in all plots
  + Cyclicity seen in first two plots
  + No trend
  + ACF only shows seasonality. Cycle length too long to show up here.

```{r, echo=TRUE, cache=TRUE}
#usdeaths
# to do
help(usdeaths)
autoplot(usdeaths)
ggseasonplot(usdeaths)
ggsubseriesplot(usdeaths)
gglagplot(usdeaths)
ggAcf(usdeaths)
```

Your answer:
+ seasonality evidend in all plots 

+ not enough data to see a cyclicity

+ no trend visible

+ maybe because there are more people outside for the summer months there are happening more accidents which lead to higher death rates





```{r, echo=TRUE, cache=TRUE}
#bricksq
# to do
help(bricksq)
autoplot(bricksq)
ggseasonplot(bricksq)
ggsubseriesplot(bricksq)
gglagplot(bricksq)
ggAcf(bricksq)
```

Your answer:
+ seasonality visible in subseriesplot and Acf plot. First and last Quarter is a little lower.

+ trend is visible in autoplot and subseriesplot. The first years the brick production is growing.

+cylicity visible in autoplot after the years 1975. There are two big decreases in clay brick production which have the same form.



```{r, echo=TRUE, cache=TRUE}
#sunspotarea
# to do
help(sunspotarea)
autoplot(sunspotarea)
gglagplot(sunspotarea, lags=20)
ggAcf(sunspotarea, lag.max = 70)

```

Your answer:
season and subseriesplot not possible because there is only one value per year.

+ seasonality visible, approximately every 10 years there is a same behaviour. PLot 1 and 3

+ trend visible in first plot. sunspotarea increases until 1950 and decreases afterwards.

+ no cyclicity visible.

+ SunspotArea has a physically explained seasonality which is visible in these plots

```{r, echo=TRUE, cache=TRUE}
#gasoline
# to do
help(gasoline)
autoplot(gasoline)
ggseasonplot(gasoline)
gglagplot(gasoline)
ggAcf(gasoline)
```

Your answer:
+ seasonality visible in acf plot. half a year the gasoline supply goes down and the other half up again.

+trend visible in first plot, gasoline supply goes up, shortly down and up again.

+cyclicity is not visible 

+ probably because the count of cars increased in the years from 1990 to 2010 very much, the supply for gasoline increased aswell. Since 2010 there a probably also many electric cars on the street which doesnt need gasoline anymore.


# Exercise 6

The `pigs` data shows the monthly total number of pigs slaughtered in Victoria, Australia, from Jan 1980 to Aug 1995. Use `mypigs <- window(pigs, start=1990)` to select the data starting from 1990. Use `autoplot` and `ggAcf` for `mypigs` series and compare these to white noise plots.

```{r, echo=TRUE, cache=TRUE}
# White noise
set.seed(1711)
y <- ts(rnorm(50))
```

```{r}
mypigs <- window(pigs, start=1990)
autoplot(y) + ggtitle("White noise")
autoplot(mypigs) + ggtitle("Pigs slaughtered")
```


```{r, echo=TRUE, cache=TRUE}
# todo
ggAcf(y)
ggAcf(mypigs)
```

Your answer:
+ In the autoplot you can see a little trend, the ammount of the pigs slaughtered over the years increase a little bit.

+ in the Acf plot one can see that the first three values are over the line that indicates if a series is a whitenoise series.
Therefore the mypigs series is no whitenoise.

# Exercise 7

`dj` contains 292 consecutive trading days of the Dow Jones Index. Use `ddj <- diff(dj)` to compute the daily changes in the index. Plot `ddj` and its ACF. Do the changes in the Dow Jones Index look like white noise?

```{r}
# todo
ddj <- diff(dj)
autoplot(ddj)
ggAcf(ddj)

```

Your answer:
+ In the autoplot i looks pretty much like white noise, because there are many seemingly unrelated spikes.

+ if you take a look at the acf you can see that this assumption is not without a reason. The most of the legs are under the indicating line for white noise. Only the 6th leg shows, that this is not only white noise.

